A meta-analysis of the impacts of best management practices on nonpoint source pollutant concentration.

### 2024-02-07 frontiers_submission/frontiers_submission.tex
wrap up discussion

\begin{abstract}
Abstract length and content varies depending on article type. Refer to
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
and length according to article type.

@@

In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{benhamLessonsLearnedTMDL2008, nationalresearchcouncilAssessingTMDLApproach2001, schrammTotalMaximumDaily2022}.
Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
The impacts of land use change on hydrology and water quality are well established \citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41\% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus \citep{epaNationalWaterQuality2017}.
FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States \citep{epaNationalWaterQuality2017}.

@@

We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
Prior meta-analysis have utilized data reported in the International Stormwater Database, which consists of self-reported and quality checked BMP data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, horvathEffectsRegionalClimate2023}.
The International Stormwater Database only recently added agricultural BMPs and has relatively sparse FIB data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014}.
Since we had interest in both FIB performance and agricultural BMPs we chose to utilize a systematic review.
We also assume that most published reports undergo some method of internal or external peer review prior to publication, which is not necessarily the case with all of the data published in the International Stormwater Database.

The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines {[}\citet{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}; Figures S1-S2{]}.
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A\&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
Fecal indicator bacteria study seraches included the following query: ``fecal indicator bacteria'' OR ``\emph{E. coli}'' OR ``\emph{Escherichia coli}'' OR ``enterococci'' OR ``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''. Nutrient BMP studies utilized a similar query: ``nutrient'' OR ``nitrogen'' OR ``phosphorus'' OR ``sediment'' OR ``TSS'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (\emph{n} = 4) was assigned a subset of studies to evaluate if they should be included (Table \ref{tab:criteria}).
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction (Table \ref{tab:datvars}), again with at least two team members reviewing each study.
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.
Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.

@@

\caption{\label{tab:modelaic}Model selection AICc.}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{AICc} \\
\cmidrule(l{3pt}r{3pt}){2-7}
Candidate Models & FIB & TN & DIN & TP & PO\textsubscript4 & TSS\\
\midrule
$\sim$ Int & 208.9 & \textbf{37.7} & \textbf{112.6} & 96.7 & 42.1 & \textbf{103.9}\\
$\sim$ Int+Influent & 197.4 & 39.8 & 114.6 & \textbf{96.3} & \textbf{36.8} & 106.4\\
$\sim$ Int+AI & 209 & 40.3 & 114.8 & 98.2 & 45.2 & 106.5\\
$\sim$ Int+AI+Influent & 195 & 42.5 & 116.9 & 98.9 & 40.2 & 109.4\\
$\sim$ Int+BMP & 209.1 & 47.3 & 121 & 101.8 & 51.8 & 115.3\\
\addlinespace
$\sim$ Int+BMP+Influent & 196.7 & 48.9 & 124 & 102.3 & 45.6 & 120.4\\
$\sim$ Int+AI+BMP & 210.1 & 50.8 & 124.1 & 104.8 & 56.4 & 120.7\\
$\sim$ Int+AI+BMP+Influent & 195.6 & 52.9 & 127.3 & 105.6 & 50.8 & 126.6\\
$\sim$ Int+BMP\texttimes Influent & 201.3 & 67.5 & 133.7 & 121.5 & 62.9 & 174.1\\
$\sim$ Int+AI\texttimes BMP & 208.1 & 75.1 & 137.6 & 121 & 67.7 & 175.5\\
\addlinespace
$\sim$ Int+Influent+AI\texttimes BMP & \textbf{193.7} & 77.2 & 141.4 & 121.6 & 70.7 & 191.4\\
$\sim$ Int+AI+BMP\texttimes Influent & 202 & 83.2 & 145.9 & 127.3 & 72.9 & 191.9\\
$\sim$ Int+AI\texttimes BMP+BMP\texttimes Influent & 201.9 & 116.8 & 156.2 & 142.7 & 94.8 & 188.5\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}Int = intercept; AI = ariditiy index; BMP = BMP subcategory}\\
\end{tabular}
\end{table}

@@

Study design prevented us from properly assessing BMP effectiveness as a function of age.
Conducting a meta-analysis of BMP effectivness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change \citep{haileHydraulicPerformancePollutant2016} which presents statistical problems, espcially when sampling variance is not reported.
Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable \citep{mitschCreatingWetlandsPrimary2012, mitschValidationEcosystemServices2014, pausAssessmentHydraulicToxic2014}.
Slopes are a valid effect size for use in meta-analysis but the set of covairates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
Future efforts for assessing the performance of BMPs over time would benefit not only from more studies, but a more standardized method for providing comparable results.

Meta-analysis indicated that BMPs resulted in significant overall reductions in FIB, TN, TP, and TSS concentrations.
We did not find strong evidence of leaching or reductions of DIN or PO\textsubscript{4} across BMP studies.
The results are in general agreement with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.
The FIB results are useful in particular because the general consensus has been that the effectiveness of BMPs at reducing FIBs and pathogens has been sparsely reviewed and understudied \citep{hagerPerformanceLowimpactDevelopment2019}.
The FIB reductions generally agreed with our hypothesis that BMP type, influent concentration, and aridity moderate the effectiveness.
The lower predicted performance in more arid regions comes with the caveat that data coverage in arid regions was quite poor, in particular for detention type BMPs.
Despite this, the results are promising considering the major limitations of using FIB as a water quality criteria.
For example, FIB can originate from non-human source, naturalize in soils, and have considerably different fate and transport mechanisms than the human pathogens of underlying concern.

The effect of higher FIB removal rates at high concentrations is expected as filtration and other removal processes associated with BMPs become saturated(?)\ldots{} something about sediment associated fib\ldots{}
The impact of aridity might be due to differential fate and transport processes in arid versus humid environments.
On one hand, we assume that arid conditions might be less hospitable to FIBs due to increased UV exposure and osmotic stress.
Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media \citep{zhangCaptureDestructionEscherichia2010, burtchettMicrozooplanktonGrazingMeans2017, deanMetaanalysisAddressingImplications2022}.
Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in these fate and transport mechanisms.
While our models capture some of the variance due to these differences as \emph{between study} effects, including these as moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.

Although we anticipated increases in nitrogen removal rates with increases in influent concentration, we did not find evidence to support this.
Increased flow rates, which can reduce residence time and increase BMP flushing, lowers nitrogen retention \citep{wollheimRetentionUrbanizingHeadwater2005, craigStreamRestorationStrategies2008}.
High nitrogen influent concentrations might be associated with higher flows and decresed BMP retention times in the included studies.
However, we did not collect associated flow data or discern between flow-weighted and mean concentration data within this study.
Many of the reviewed studies appear to fail to include associated flow volume information.

We also did not find evidence that BMP type or aridity moderated nitrogen removal.
This is result is largely consistent with findings in reviews by \citet{kochNitrogenRemovalStormwater2014}, \citet{hagerPerformanceLowimpactDevelopment2019} and \citet{horvathEffectsRegionalClimate2023}.
There are a large number of abiotic and biotic processes that control nitrogen retention and removal in BMPs and these processes are moderated by both site specific climate and design factors \citep{lefevreReviewDissolvedPollutants2015, valencaNitrateRemovalUncertainty2021}.
It is likely that these site specific factors (retained soil moisture, submerged anoxic zones, vegetation, media composition) are not captured by our broad categorization of BMP types and aridity index values.
For example, \citet{valencaNitrateRemovalUncertainty2021}, using data from the International Stormwater Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.

\emph{compare nutrient effects to koch, barrett, clary}

\emph{Linking BMP performance to watershed scale improvement remains a challenge (Tomer and Locke, Melland et al, Meals).}
\emph{The succesful implementation and maintence of BMPs at watershed scales remains an area for more investigation (Liu et al., Lintern).}
\emph{Long-term trans-disciplinary monitoring projects not just on BMP performance but maintance, management, and social apsects may elucidate unknown factors influencing the ability to scale BMP reductions to watershed scale improvements (Lintern et al.)}

\emph{Runoff volume was not incorporated into our models, largely attributed to lack of available data. However, BMP size is typically scaled to accommodate runoff volumes of a particular return period volume (1-yr or 2-yr flows for example) (Hoss et al.).}
\emph{data exploration did not suggest this was the case, but it is reasonable to suggest the the underlying relationship between ROM and any of the moderators are a nonlinear function. Semi-parametric approaches might provide some additional insight}

>>>>

\begin{abstract}
We conducted a systematic review of published nonpoint source water quality best management practice (BMP) studies and used a meta-analysis approach to describe variance in pollutant removal performance. Despite high variance, we found the BMPs on average were effective at reducing fecal indicator bacteria (FIB), total nitrogen (TN), total phosphorus (TP), and total suspended sediment (TSS) concentrations. We explored how much BMP pollutant removal process, influent pollutant concentration, and aridity effected BMP performance. We found that influent concentration and interaction effect between the BMP pollutant removal process and aridity explained a substantial amount of variance in BMP performance in FIB removal. Influent concentration explained a small amount of variability in BMP removal of TP and orthophosphate (PO\textsubscript{4}). We did not find evidence that any of our chosen variables moderated BMP performance in nitrogen or TSS removal. Through our systematic review, we found inadequate spatial representation of BMP studies to capture the underlying variability in climate, soil, and other conditions that could impact BMP performance.

@@

In the United States (U.S.), major improvement in water quality has been achieved under the Clean Water Act of 1972.
This progress has been largely attributed to investments and reductions in point source discharges while reductions in nonpoint source pollutants remain a substantial challenge \citep{nationalresearchcouncilAssessingTMDLApproach2001, benhamLessonsLearnedTMDL2008, schrammTotalMaximumDaily2022}.
Increased pollutant loads and concentrations in runoff resulting from land use changes are a particular challenge.
The impacts of land use change on hydrology and water quality are well established \citep{allanLandscapesRiverscapesInfluence2004, carpenterNonpointPollutionSurface1998, bernhardtUnderstandingManagingMinimizing2008, careyEvaluatingNutrientImpacts2013, freemanImpactsUrbanizationDevelopment2019}.
Nonpoint source driven fecal indicator bacteria (FIB), nitrogen, phosphorus, and suspended sediment remain major causes of water quality impairments in U.S. rivers and streams despite decades of work.
In 2017, the Environmental Protection Agency (EPA) estimated 41\% or more of the nation's rivers and streams rated poorly for biological condition due to excess nitrogen or phosphorus \citep{epaNationalWaterQuality2017}.
FIB remains the leading cause of water body impairment on the Clean Water Act 303(d) list in the United States \citep{epaNationalWaterQuality2017}.

@@

We conducted a systematic review of recent (2000-2022) literature to compile field studies documenting the effectiveness of best management practices on fecal indicator bacteria and nutrients concentrations.
Prior meta-analysis have utilized data reported in the International Stormwater BMP Database (\url{https://bmpdatabase.org/}), which consists of self-reported and quality checked BMP data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, horvathEffectsRegionalClimate2023}.
The International Stormwater BMP Database only recently added agricultural BMPs and has relatively sparse FIB data \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014}.
Since we had interest in both FIB performance and agricultural BMPs we chose to utilize a systematic review.

The systematic review followed guidance provided in the Collaboration for Environmental Evidence systematic review guidelines \citep{collaborationforenvironmentalevidenceGuidelinesStandardsEvidence2018}.
In order to maximize the number of studies included in the review, we included both peer-reviewed studies and unpublished white papers to reduce potential bias against negative results.
The inclusion criteria filtered out (1) non-field studies, (2) modelling results, (3) studies that did not evaluate specific BMPs, (4) studies conducted outside of the U.S. or published in a language other than English.
We ran search queries in Texas A\&M Library Catalog, Web of Science, and Google Scholar.
Although results from Google Scholar are not always replicable, we utilized the service to maximize search results for studies not published in academic journals and presumably increase the chance of identifying studies with negative effects.
Fecal indicator bacteria study searches included the following query: ``fecal indicator bacteria'' OR ``\emph{E. coli}'' OR ``\emph{Escherichia coli}'' OR ``enterococci'' OR ``enterococcus'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''. Nutrient BMP studies utilized a similar query: ``nutrient'' OR ``nitrogen'' OR ``phosphorus'' OR ``sediment'' OR ``TSS'' AND ``best management practices'' OR ``BMPs'' AND ``effectiveness'' OR ``performance''.

Results from each database were first filtered to remove duplicates.
After removal of duplicates, each member of the research team (\emph{n} = 4) was assigned a subset of studies to evaluate if they should be included (Table \ref{tab:criteria}).
Each study was reviewed by two team members and differences in opinion were collectively discussed and agreed upon before progressing.
The remaining studies were split among team members for data extraction (Table \ref{tab:datvars}), again with at least two team members reviewing each study.
If data was provided in figures, the data was extracted with the WebPlotDigitizer tool \citep{rohatgiWebPlotDigitizer2022}.
Searches, review, and data extractions were conducted separately for FIB and nutrient related parameters.
See Figure S1 and S2 for ROSES diagram.

@@

\caption{\label{tab:modelaic}Sumary of AICc values used for model selection. Bolded values indicate the selected candidate model.}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{AICc} \\
\cmidrule(l{3pt}r{3pt}){2-7}
Candidate Models & FIB & TN & DIN & TP & PO\textsubscript4 & TSS\\
\midrule
$\sim$ Int & 208.9 & \textbf{37.7} & \textbf{112.6} & 96.7 & 42.1 & \textbf{103.9}\\
$\sim$ Int+Influent & 197.4 & 39.8 & 114.6 & \textbf{96.3} & \textbf{36.8} & 106.4\\
$\sim$ Int+AI & 209 & 40.3 & 114.8 & 98.2 & 45.2 & 106.5\\
$\sim$ Int+AI+Influent & 195 & 42.5 & 116.9 & 98.9 & 40.2 & 109.4\\
$\sim$ Int+BMP & 209.1 & 47.3 & 121 & 101.8 & 51.8 & 115.3\\
\addlinespace
$\sim$ Int+BMP+Influent & 196.7 & 48.9 & 124 & 102.3 & 45.6 & 120.4\\
$\sim$ Int+AI+BMP & 210.1 & 50.8 & 124.1 & 104.8 & 56.4 & 120.7\\
$\sim$ Int+AI+BMP+Influent & 195.6 & 52.9 & 127.3 & 105.6 & 50.8 & 126.6\\
$\sim$ Int+BMP\texttimes Influent & 201.3 & 67.5 & 133.7 & 121.5 & 62.9 & 174.1\\
$\sim$ Int+AI\texttimes BMP & 208.1 & 75.1 & 137.6 & 121 & 67.7 & 175.5\\
\addlinespace
$\sim$ Int+Influent+AI\texttimes BMP & \textbf{193.7} & 77.2 & 141.4 & 121.6 & 70.7 & 191.4\\
$\sim$ Int+AI+BMP\texttimes Influent & 202 & 83.2 & 145.9 & 127.3 & 72.9 & 191.9\\
$\sim$ Int+AI\texttimes BMP+BMP\texttimes Influent & 201.9 & 116.8 & 156.2 & 142.7 & 94.8 & 188.5\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}Int = intercept; AI = ariditiy index; BMP = BMP subcategory}\\
\end{tabular}
\end{table}

@@

Study design prevented us from properly assessing BMP effectiveness as a function of age.
Conducting a meta-analysis of BMP effectivness over time is hampered both by the lack of long-term studies and lack of standardized reporting mechanisms.
Some studies simply describe the change in pollutant concentrations or loads at the beginning and end of the study as a percent change \citep{haileHydraulicPerformancePollutant2016} which presents statistical problems, espcially when sampling variance is not reported.
Changes in performance can also be described using a linear regression using date (transformed as a numeric variable) as an independent variable and log-transformed water quality as the dependent variable \citep{mitschCreatingWetlandsPrimary2012, mitschValidationEcosystemServices2014, pausAssessmentHydraulicToxic2014}.
Slopes are a valid effect size for use in meta-analysis but the set of covariates used between studies should be the same since the coeffecient of interest is adjusted to account for other terms in the the regression model \citep{beckerSynthesisRegressionSlopes2007}.
It would be reasonable to assume that regressions equations vary between studies to adjust results for seasonality, flow rates, and other variables.
Future efforts for assessing the performance of BMPs over time would benefit not only from more studies, but a more standardized method for providing comparable results.

Meta-analysis indicated that BMPs resulted in significant overall reductions in FIB, TN, TP, and TSS concentrations.
We did not find strong evidence of leaching or reductions of DIN or PO\textsubscript{4} across BMP studies.
The results are in general agreement with previous reviews that found effective (but highly variable) removal efficiencies for nitrogen, phosphorus, and sediment \citep{claryBMPPerformanceAnalysis2011, kochNitrogenRemovalStormwater2014, liuReviewEffectivenessBest2017}.
The FIB results are useful in particular because FIBs performance by BMPs have been sparsely reviewed and generally understudied \citep{hagerPerformanceLowimpactDevelopment2019}.

The FIB reductions generally agreed with our hypothesis that BMP type, influent concentration, and aridity moderate the effectiveness.
The lower predicted performance in more arid regions comes with the caveat that data coverage in arid regions was quite poor, in particular for detention type BMPs.
Despite this, the results are promising considering the major limitations of using FIB as a water quality criteria.
It is important to note that FIB can originate from non-human source and naturalize in soils, and result in different underlying risk of illness \citep{ishiiEscherichiaColiEnvironment2008, schoenAssessingPathogenRisk2010, sollerEstimatedHumanHealth2010, fujiokaRecreationalWaterQuality2015}.
Since the fate and transport of human pathogens within BMPs can potentially differ from FIB, BMP choices probably should not be based on FIB reduction alone as alternative indicators or even direct pathogen measurement becomes available \citep{pengIndicatorPathogenRemoval2016, waltersPersistenceNucleicAcid2009}.
However, relatively few studies have compared human pathogen and FIB removal rates within BMPs \citep{rughHighlyVariableRemoval2022}.

While we observed a strong relationship between FIB influent concentration and FIB removal across all BMPs, we anticipated this relationship to vary by BMP subcategory.
The reliance on certain removal processes by BMP subcategories was expected to effect the ability of the BMP to retain FIB at higher or lower concentrations.
We did have some evidence of differing BMP subcateogy removal under different ariditiy indecies.
The impact of aridity might be due to differential fate and transport processes in arid versus humid environments.
On one hand, we assume that arid conditions might be less hospitable to FIBs due to increased UV exposure and osmotic stress.
Conversly, these conditions are also less hospitable to the protoza, bacteriophages, and micro-zooplankton that can play a strong role in predating on and controlling FIB concentrations within BMP media \citep{zhangCaptureDestructionEscherichia2010, burtchettMicrozooplanktonGrazingMeans2017, deanMetaanalysisAddressingImplications2022}.
Site specific conditions (such as retained soil moisture, turbidity, vegetation, and other factors) play an important role in bacteria survival as well as for influencing the filtration and attachment processes that retain FIB within BMP media.
For example, the presence or absence of a submerged zone within a bioretention BMP has a strong effect on FIB removal \citep{rippyMeetingCriteriaLinking2015, pengIndicatorPathogenRemoval2016}.
While our models capture some of the variance due to these differences as between study effects, including these variables as fixed effect moderators in a meta-regression model would be valuable but these details are under reported in BMP studies.

Although we anticipated increases in nitrogen removal rates with increases in influent concentration, we did not find evidence to support this.
Increased flow rates, which can reduce residence time and increase BMP flushing, lowers nitrogen retention \citep{wollheimRetentionUrbanizingHeadwater2005, craigStreamRestorationStrategies2008}.
High nitrogen influent concentrations might be associated with higher flows and decreased BMP retention times in the included studies.
However, we did not collect associated flow data or discern between flow-weighted and mean concentration data within this study.
Many of the reviewed studies appear to fail to include associated flow volume information.

We also did not find evidence that BMP type or aridity moderated nitrogen removal.
This result is largely consistent with findings in reviews by \citet{kochNitrogenRemovalStormwater2014}, \citet{hagerPerformanceLowimpactDevelopment2019} and \citet{horvathEffectsRegionalClimate2023}.
There are a large number of abiotic and biotic processes that control nitrogen retention and removal in BMPs and these processes are moderated by both site specific climate and design factors \citep{lefevreReviewDissolvedPollutants2015, valencaNitrateRemovalUncertainty2021}.
It is likely that these site specific factors (retained soil moisture, submerged anoxic zones, vegetation, media composition) are not captured by our broad categorization of BMP types and aridity index values.
For example, \citet{valencaNitrateRemovalUncertainty2021}, using data from the International Stormwater BMP Database, showed that the relative importance of climate and design variables for moderating nitrogen removal varied by BMP type.

Although influent phosphorus concentration was included in the selected TP and PO\textsubscript{4} models, they provided relatively low explanatory ability.
By comparison, \citet{horvathEffectsRegionalClimate2023} found three types of BMPs resulted in differing TP and dissolved inorganic phosphorus removal rates with influent concentrations explaining a small proportion of removal rate variance.
Again, site-specific factors not captured in our broad categorizations of BMP type and aridity index play a role in differential phosphorus removal rates.
Soil physical characteristics and media amendments (iron for example) can play an critical role in sorption capacity and are dependent on covariates such as contact time and pH \citep{hoganBestManagementPractices2007, lefevreReviewDissolvedPollutants2015}.

A few reviews have noted a common trend of insufficient methodological and site specific data among peer-reviewed BMP performance studies \citep{grudzinskiDoesRiparianFencing2020, eagleMetaAnalysisConstrained2017, liuReviewEffectivenessBest2017}.
We confirm that inconsistent reporting among studies complicates data extraction, effect size calculations, and attributing important sources of variance.
The International Stormwater BMP Database addresses some of these concerns through a standardized reporting format.
Our future efforts will incorporate data from the International Stormwater BMP Database with data retrieved through a systematic review.

In summary, we used multi-level random effects meta-regression models to estimate overall BMP effectiveness from systematically reviewed studies.
Although there was relatively high variance between studies, we found strong evidence that BMPs reduce overall mean FIB, TN, TP, and TSS concentrations.
Influent concentrations moderated BMP efficiency for both FIB and PO\textsubscript{4}, with larger removal rates at high influent concentrations.
We found that aridity and BMP subcategory moderated BMP performance for only FIB.
Most likely, site specific design and climate variables not captured in our review or by our choice in BMP classification approach play a more important role in explaining BMP performance variability.
Future efforts should seek to retrieve more detailed study information.
Furthermore, our systematic review highlights the poor spatial coverage of BMP studies.
The reviewed studies therefore fail to incorporate the range of soil, climate, and runoff conditions needed to adequately link BMP performance to local predictors.

Scaling BMP pollutant reductions to basin wide water quality improvement remains a substantial challenge \citep{tomerChallengeDocumentingWater2011, mellandEffectsAgriculturalLand2018, mealsLagTimeWater2010}.
The major identified challenges include the lack of long term studies, inadequate data collected on BMP management, and incomplete understanding of BMP function \citep{linternBestManagementPractices2020, liuReviewEffectivenessBest2017}.
There is a clear need to fill knowledge gaps through additional long-term spatially relevant BMP studies.
However, we emphasize the need for convergent research approaches that better align study design and reporting that produces data aligned with data synthesis and modeling approaches.
